---
title: "Kawasaki vs. Viral Infections"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

# Inclusion criteria
The *patients* inclusion criteria is **patients < 18 at the date of visit**
The *Visits* inclusion criteria includes:
    -   Inpatients (at least one hospitalization \> 1 day)
    -   Comprising at least one ICD-10 code during hospitalization
    -   Start date: January 1st, 2019

# Variables that need to be checked/modified by each site
Change the values of the following variables according to your site:
1.  "folder_4ce_files": folder path where your phase 2.2 data files are
    located
2.  "obfuscation": determine the obfuscation threshold (FALSE if no
    obfuscation, or the numeric value of the obfuscation threshold if
    any)
3.  "dateFormat": specify the format of the date at your site (e.g., for
    "2021-06-24", the format would be ""%Y-%m-%d"", [see
    documentation](https://www.stat.berkeley.edu/~s133/dates.html))
4.  "data_update_date": date at which the data has been updated in the
    local datawarehouse. Used to estimate patient age at time of visit,
    since patients age in the 4CE demographic file is expected the age
    at data update.


```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/4ceData/Input/4CEAllAdminChortsOct2023/"
obfuscation =  FALSE
#obfuscation = 3
dateFormat <- "%Y-%m-%d"
data_update_date <- "2023-10-01"
age_estimation_variability <- FALSE
```


# Set up
- Installing and loading the libraries
- Get the session info 

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
source("../R/packagesInstallation.R")
# get the session information
sessionInfo()
# load the local variables
source("../R/setUpVariableParameters.R")

print(paste0( 'FourCePhase2.2kawasai_ad Code version: ', codeVersion ))
print('##########')
```

# Data loading
We will use as input the 2.2 data files. Specifically:
-   LocalPatientSummary
-   LocalPatientObservation
-   LocalPatientClinicalcourse

## Reading 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files without duplicating the data

source("../R/readInputFiles_fread.R")

filecheck <- checkInputFilesFread( path = folder_4ce_files, separator = ",", skip = 0, verbose = TRUE )

if(filecheck){
  demo_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientSummary.csv"), sep = ',',skip = 0, fill=TRUE )
  obs_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientObservations.csv"), sep = ',',skip = 0, fill=TRUE)
  clinical_raw <-  data.table::fread(paste0( folder_4ce_files, "LocalPatientClinicalCourse.csv"), sep = ',',skip = 0, fill=TRUE)

  ### Extract the patient summary and observation information. 
  # arrange in a consistent way to make sure we are not assuming the order of some sites
  # select only the study period
  demo_raw <- demo_raw %>% 
    dplyr::mutate( age = trunc( age ) ) %>%
    dplyr::filter(cohort == "AllAdm", 
                  age >= 0,
                  as.Date(admission_date, format = dateFormat) >= ymd('2019-01-01'),
                  as.Date(admission_date, format = dateFormat) <= ymd('2024-04-01')) %>%
    dplyr::arrange(patient_num, as.Date(admission_date, format = dateFormat), days_since_admission)
  obs_raw <- obs_raw %>% 
    dplyr::filter(cohort == "AllAdm") %>%
    dplyr::arrange(patient_num, days_since_admission)
  clinical_raw <- clinical_raw %>% 
    dplyr::filter(cohort == "AllAdm",
                  as.Date(calendar_date, format = dateFormat) >= ymd('2019-01-01'),
                  as.Date(calendar_date, format = dateFormat) <= ymd('2024-04-01')) %>%
    dplyr::arrange(patient_num, as.Date(calendar_date, format = dateFormat), days_since_admission)
  
  print('Input files successfully read, patient data extracted')
  
}else{
  print('Error in reading files')
}

#remove potential duplicates
clinical_raw <- clinical_raw %>% unique()
demo_raw <- demo_raw %>% unique()
obs_raw <- obs_raw %>% unique()

```

## Select the ICD codes of interest
```{r}
print('Reading ICD code')
kawasaki_Codes <- read.csv("../public-data/ICD_kawasaki.csv", header = TRUE, colClasses = "character")
viral_infection <- read.csv("../public-data/phecode_icd10.csv") %>%
    dplyr::mutate( concept_code = gsub("[.]", "", ICD10)) %>%
    dplyr::select( ICD10_Code = concept_code, description = `ICD10.String`, category = Phenotype) %>%
    dplyr::filter( category == "Viral infection")

icdCodes <- rbind( kawasaki_Codes, viral_infection )
```

# Data-management

## Filtering on "inpatient" status

```{r}
# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
clinical_raw <- dplyr::filter(clinical_raw, in_hospital == 1)
```

## Calculating age at hospitalization time

```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
# removing the dots from the ICD codes (if any)
obs_raw <- as.data.frame( obs_raw )
demo_raw <- as.data.frame( demo_raw )
clinical_raw <- as.data.frame( clinical_raw )

obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age", "dead", "death_date")], by = c("patient_num")) %>%
  dplyr::mutate( concept_code = gsub("\\.", "",concept_code )) %>%
  dplyr::inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))

# estimate age at each diagnosis
if( age_estimation_variability == FALSE ){
  
  obs_raw_age <- obs_raw_age %>%
    dplyr::mutate( age_time_diagnosis = age - floor(as.numeric(as.Date(data_update_date, format = dateFormat) - as.Date(calendar_date, format = dateFormat)) / 365.25)) %>%
    dplyr::select( -dead, -death_date )
  
}else if( age_estimation_variability == TRUE ){
  
  alive<- obs_raw_age %>%
    dplyr::filter( dead == 0 ) %>%
    dplyr::mutate( age_time_diagnosis = age - floor(as.numeric(as.Date(data_update_date, format = dateFormat) - as.Date(calendar_date, format = dateFormat)) / 365.25 )) %>%
    dplyr::select( -dead, -death_date )
  
  dead<- obs_raw_age %>%
    dplyr::filter( dead == 1 ) %>%
    dplyr::mutate( age_time_diagnosis = age - floor(as.numeric(as.Date(death_date, format = dateFormat) - as.Date(calendar_date, format = dateFormat)) / 365.25 )) %>%
    dplyr::select( -dead, -death_date )
 obs_raw_age <- rbind( alive, dead )
}

print('Patient age calculated')
```

## Applying inclusion criteria to filter and format the data

We are selecting hospitalizations based on the following criteria:

-   Associated with at least 1 ICD code

-   With a duration of hospitalization spanning over at least one
    consecutive days

-   Starting on or after 2019/01/01

-   Age at diagnosis < 18 y/o

```{r}
obs_filtered_inclusion_criteria <- obs_raw_age %>% 
  dplyr::filter(concept_type == 'DIAG-ICD10',
                days_since_admission >= 0, 
                age_time_diagnosis < 18 )
```

## Select the disorders of interest based on the ICD code
```{r}
obs_disease_interest <- obs_filtered_inclusion_criteria %>%
  dplyr::filter( concept_code %in% icdCodes$ICD10_Code ) 
```

## Estimate the length of the hospitalization
```{r}
source("../R/count_sequences_hospitalisation.R")

clinical_raw_doi <- clinical_raw %>%
  dplyr::filter( patient_num %in% obs_disease_interest$patient_num )

clinical_raw_doi <- count_hosp(clinical_raw_doi) %>%
  dplyr::ungroup() %>%
  dplyr::select( patient_num, days_since_admission, n_hospitalization, length_hospitalization, admit_date )

```

## Hospitalizations per month and disease subtype to plot 
```{r}
toPlot_formated <- obs_disease_interest %>%
  dplyr::left_join( clinical_raw_doi, by=c("patient_num", "days_since_admission")) %>%
  dplyr::mutate( time_p = as.Date(cut( as.Date(admit_date), breaks = "months"))) %>%
  dplyr::filter( as.Date(time_p) >= ymd(start_date_plots),
                 as.Date(time_p) <= ymd(end_date_plots)) %>%
  dplyr::left_join( icdCodes, by=c("concept_code"="ICD10_Code"))
  
toPlot_counts <- toPlot_formated %>%
  dplyr::mutate( pat_hosp = paste0( patient_num, "-", n_hospitalization )) %>%
  dplyr::group_by( time_p, category ) %>%
  #dplyr::summarise( n = n_distinct( pat_hosp )) %>%
  dplyr::summarise( n = ifelse(n_distinct(pat_hosp) > obfuscation | isFALSE(obfuscation),
                               n_distinct(pat_hosp), 
                               0.5))
```

## Plot cases per month
```{r}
toPlotTwoAxis <- toPlot_counts %>%
  tidyr::pivot_wider( names_from = category, values_from = n )

ggplot(data = toPlotTwoAxis, aes(x = time_p)) +
  geom_line(aes(y = `Kawasaki disease`, colour="Kawasaki Disease", linetype = "Kawasaki Disease")) +
  geom_line(aes(y = `Viral infection` / 20, colour="Viral Infections", linetype = "Viral Infections" )) + 
  scale_y_continuous(
    name = "Kawasaki Cases",
    sec.axis = sec_axis(~ . * 20, name = "Viral Cases")  # Scale up for secondary axis
  ) +
  scale_color_manual(
    name = "Legend",
    values = c("Kawasaki Disease" = "orange", "Viral Infections" = "darkblue")
  ) +
  scale_linetype_manual(
    name = "Legend",
    values = c("Kawasaki Disease" = "dashed", "Viral Infections" = "solid")
  ) +
  labs(x = "Time Period") +
  theme_minimal() +
  theme( legend.position = "bottom")

```

